### 2023-11-24 scripts/01_sim_data.qmd
simulate data1

---
title: "01 - Simulate data"
format: html
---

```{r}
#| label: setup

library(tidyverse)
theme_set(theme_light())
library(mgcv)
my_seed <- 4785
```

## Data 1: U-shaped effect

We simulate data of the U-shaped effect of language acquisition. Subjects take a test at 11 time points where they language proficiency is measured and a score value is obtained for how well they have learnt construction C.

```{r}
#| label: dat1

set.seed(my_seed)

subjs <- 200
proficiency <- runif(subjs, -2, 2)
proficiency_b <- rnorm(subjs, 0.1, 0.05)
timep <- 11
intercept <- 15

proficiency_subjs <- list()
score_subjs <- list()

for (subj in 1:subjs) {
  # Get proficiency for subject `subj` in time point 1 to 11
  proficiency_subj <- proficiency[subj] + proficiency_b[subj] * 0:10
  # Calculate score of subject `subj` based on their proficiency
  score_subj <- intercept + proficiency_subj^2 + rnorm(1, 0, 0.5)
  # Add random noise to score
  score_subj <- score_subj * runif(1, 1, 2) + rnorm(timep, 0, 0.1)

  # Add proficiency values to list
  proficiency_subjs[subj] <- list(proficiency_subj)
  # Add scores to list
  score_subjs[subj] <- list(score_subj)
}

dat1 <- tibble(
  score = unlist(score_subjs),
  proficiency = unlist(proficiency_subjs),
  subj = as.factor(paste("s", rep(1:subjs, each = timep), sep = "")),
  time = rep(0:10, length.out = subjs * timep)
)
```

```{r}
#| label: dat1-pro-sco

dat1 %>%
  ggplot(aes(proficiency, score, group = subj)) +
  geom_path(alpha = 0.5) +
  geom_point(alpha = 0.5)
```

```{r}
#| label: dat1-pro-sco-t5

dat1 %>%
  filter(time == 5) %>%
  ggplot(aes(proficiency, score, group = subj)) +
  geom_point(alpha = 0.5)
```

```{r}
#| label: dat1-subj-12
set.seed(my_seed)

dat1 %>%
  filter(subj %in% sample(unique(dat1$subj), 12)) %>%
  ggplot(aes(time, score, group = subj, colour = proficiency)) +
  geom_path() +
  geom_point(alpha = 0.5) +
  facet_wrap(vars(subj))
```

```{r}
#| label: dat1-subj-12-pro-sco
set.seed(my_seed)

dat1 %>%
  filter(subj %in% sample(unique(dat1$subj), 12)) %>%
  ggplot(aes(proficiency, score, group = subj)) +
  geom_path(alpha = 0.5) +
  geom_point(alpha = 0.5)
```

## Data 2:

We simulate word-final vowel reduction in learners of Spanish.

- /i/, /u/, /a/
  - i: bikini, casi, kiwi, cadi, pony, taxi, yeti, cursi, caqui, ferry
  - u: tofu, tribu, urdu, haiku, ecu, quipu, 
  - a: casa, pasa, taza, masa, para, cosa, duda, tema, ropa, bella

Avg. formant values for monolingual English and Spanish speakers

|         |     | /i/        | /a/       | /u/       |
| :------ | :-- | ---------: | --------: | --------: |
| Spanish | F1  | 286 (6)    | 638 (36)  | 322 (20)  |
|         | F2  | 2147 (131) | 1353 (84) | 992 (121) |
|         |     |            |           |           |
| English | F1  | 264 (34)   | 783 (155) | 316 (43)  |
|         | F2  | 2268 (207) | 1182 (152)| 1183 (153)|