### 2023-11-27 scripts/01_sim_data.qmd
edit dat2

---
title: "01 - Simulate data"
format: html
---

```{r}
#| label: setup

library(tidyverse)
theme_set(theme_light())
library(mgcv)
my_seed <- 4785
```

## Data 1: U-shaped effect

We simulate data of the U-shaped effect of language acquisition. Subjects take a test at 11 time points where they language proficiency is measured and a score value is obtained for how well they have learnt construction C.

```{r}
#| label: dat1

set.seed(my_seed)

subjs <- 200
proficiency <- runif(subjs, -2, 2)
proficiency_b <- rnorm(subjs, 0.1, 0.05)
timep <- 11
intercept <- 15

proficiency_subjs <- list()
score_subjs <- list()

for (subj in 1:subjs) {
  # Get proficiency for subject `subj` in time point 1 to 11
  proficiency_subj <- proficiency[subj] + proficiency_b[subj] * 0:10
  # Calculate score of subject `subj` based on their proficiency
  score_subj <- intercept + proficiency_subj^2 + rnorm(1, 0, 0.5)
  # Add random noise to score
  score_subj <- score_subj * runif(1, 1, 2) + rnorm(timep, 0, 0.1)

  # Add proficiency values to list
  proficiency_subjs[subj] <- list(proficiency_subj)
  # Add scores to list
  score_subjs[subj] <- list(score_subj)
}

dat1 <- tibble(
  score = unlist(score_subjs),
  proficiency = unlist(proficiency_subjs),
  subj = as.factor(paste("s", rep(1:subjs, each = timep), sep = "")),
  time = rep(0:10, length.out = subjs * timep)
)
```

```{r}
#| label: dat1-pro-sco

dat1 %>%
  ggplot(aes(proficiency, score, group = subj)) +
  geom_path(alpha = 0.5) +
  geom_point(alpha = 0.5)
```

```{r}
#| label: dat1-pro-sco-t5

dat1 %>%
  filter(time == 5) %>%
  ggplot(aes(proficiency, score, group = subj)) +
  geom_point(alpha = 0.5)
```

```{r}
#| label: dat1-subj-12
set.seed(my_seed)

dat1 %>%
  filter(subj %in% sample(unique(dat1$subj), 12)) %>%
  ggplot(aes(time, score, group = subj, colour = proficiency)) +
  geom_path() +
  geom_point(alpha = 0.5) +
  facet_wrap(vars(subj))
```

```{r}
#| label: dat1-subj-12-pro-sco
set.seed(my_seed)

dat1 %>%
  filter(subj %in% sample(unique(dat1$subj), 12)) %>%
  ggplot(aes(proficiency, score, group = subj)) +
  geom_path(alpha = 0.5) +
  geom_point(alpha = 0.5)
```

## Data 2:

We simulate word-final vowel reduction in learners of Spanish.

- /i/, /u/, /a/
  - i: bikini, casi, kiwi, cadi, pony, taxi, yeti, cursi, caqui, ferry
  - u: tofu, tribu, urdu, haiku, ecu, quipu, 
  - a: casa, pasa, taza, masa, para, cosa, duda, tema, ropa, bella

Avg. formant values for monolingual English and Spanish speakers

|         |     | /i/        | /a/       | /u/       |
| :------ | :-- | ---------: | --------: | --------: |
| Spanish | F1  | 286 (6)    | 638 (36)  | 322 (20)  |
|         | F2  | 2147 (131) | 1353 (84) | 992 (121) |
|         |     |            |           |           |
| English | F1  | 264 (34)   | 783 (155) | 316 (43)  |
|         | F2  | 2268 (207) | 1182 (152)| 1183 (153)|

```{r}
#| label: eud
# https://mycurvefit.com

x <- 1:9

y_esi <- rep(2.25, 9)
y_esu <- rep(2, 9)
y_esa <- rep(1.5, 9)

y_es0i <- 2 + (-0.3 - 2)/(1 + (x/2)^2.5)
y_es0u <- 1.6 + (-4 - 1.6)/(1 + (x/0.5)^1.5)
y_es0a <- rep(0, 9)

y_es1i <- 2 + (0.9 - 2)/(1 + (x/2)^2.5)
y_es1u <- 1.5 + (1 - 1.5)/(1 + (x/2)^2.5)
y_es1a <- rep(1.25, 9)
```

```{r}
#| label: eud-plot

tibble(
  time = rep(x, 9),
  eud = c(y_esi, y_esu, y_esa, y_es0i, y_es0u, y_es0a, y_es1i, y_es1u, y_es1a),
  group = rep(c("es", "es_0", "es_1"), each = 9*3),
  vowel = rep(rep(c("i", "u", "a"), each = 9), 3)
) %>%
  ggplot(aes(time, eud, colour = vowel)) +
  geom_point(size = 3) +
  facet_grid(cols = vars(group))
```

```{r}
set.seed(my_seed)

subjs <- 60
timep <- 9
reps <- 6

dat2_grid <- expand_grid(
  subj = 1:subjs,
  group = c("es", "es_0", "es_1"),
  vowel = c("i", "u", "a"),
  rep = 1:reps
)

subj_int <- rnorm(subjs * 3, 0, 0.15)
subj_slo <- rnorm(subjs * 3, 0, 0.25)

dat2 <- tibble()

for (rown in 1:nrow(dat2_grid)) {
  # Get current row, group and vowel
  this_row <- dat2_grid[rown,]
  this_subj <- this_row$subj
  this_group <- this_row$group
  this_vowel <- this_row$vowel
  this_int <- rnorm(1, 0, 0.2)
  this_slo <- rnorm(1, 0, 0.2)

  # Calculate EUD trajectory
  if (this_group == "es") {
    if (this_vowel == "i") {
      this_eud <- rep(2.25 + subj_int[this_subj] + this_int, 9)
    } else if (this_vowel == "u") {
      this_eud <- rep(2 + subj_int[this_subj] + this_int, 9)
    } else if (this_vowel == "a") {
      this_eud <- rep(1.5 + subj_int[this_subj] + this_int, 9)
    }

  } else if (this_group == "es_0") {
    if (this_vowel == "i") {
      this_eud <- 2 + (-0.3 + subj_int[this_subj] + this_int - 2) /
        (1 + (1:timep/2)^(2.5 + (subj_slo[this_subj] + this_slo) * 2))
    } else if (this_vowel == "u") {
      this_eud <- 1.6 + (-4 + subj_int[this_subj] + this_int - 1.6) /
        (1 + (1:timep/0.5)^(1.5 + subj_slo[this_subj]  + this_slo))
    } else if (this_vowel == "a") {
      this_eud <- rep(0 + subj_int[this_subj] + this_int, 9)
    }

  } else if (this_group == "es_1") {
    if (this_vowel == "i") {
      this_eud <- 2 + (0.9 + subj_int[this_subj] + this_int - 2) /
        (1 + (1:timep/2)^(2.5 + subj_slo[this_subj] + this_slo))
    } else if (this_vowel == "u") {
      this_eud <- 1.5 + (1 + subj_int[this_subj] + this_int - 1.5) /
        (1 + (1:timep/0.5)^(2.5 + subj_slo[this_subj] + this_slo))
    } else if (this_vowel == "a") {
      this_eud <- rep(1.25 + subj_int[this_subj] + this_int, 9)
    }
  }

  # Add eud to dat2
  this_row$eud <- list(this_eud)
  this_trial <- unnest(this_row, eud)
  this_trial$timep <- 1:timep
  dat2 <- bind_rows(dat2, this_trial)
}

dat2 <- dat2 %>%
  mutate(
    id = paste(group, subj, rep, sep = ".")
  )
```

```{r}
dat2 %>%
  ggplot(aes(timep, eud, group = id, colour = group)) +
  geom_path(alpha = 0.15, linewidth = 1) +
  facet_grid(cols = vars(vowel), rows = vars(group))
```

```{r}
dat2 %>%
  ggplot(aes(timep, eud, colour = group)) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 3)) +
  facet_grid(cols = vars(vowel), rows = vars(group))
```

```{r}
dat2 %>%
  filter(subj == 1, group == "es_0") %>%
  ggplot(aes(timep, eud, group = id, colour = as.factor(rep))) +
  geom_path() +
  facet_grid(cols = vars(vowel))
```