### 2023-12-02 scripts/02_data_analysis.qmd

---
title: "02 Data analysis"
format: html
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
theme_set(theme_light())
library(mgcv)
library(tidygam)
library(ggeffects)
my_seed <- 4785
```

## Read data

```{r}
dat1 <- readRDS("data/dat1.rds")
dat2 <- readRDS("data/dat2.rds")
```

## Data 1

### Score and proficiency (single time point)

Let's start with data from a single time point and let's model score as a function of proficiency.

```{r}
#| label: dat1-5

dat1_5 <- dat1 %>%
  filter(time == 5)

dat1_5
```

There are `r length(unique(dat1_5$subj))` subjects. The median score is `r median(dat1_5$score)` with a range between `r min(dat1_5$score)` and `r max(dat1_5$score)`. The median proficiency is `r median(dat1_5$proficiency)` with a range between `r min(dat1_5$proficiency)` and `r max(dat1_5$proficiency)`.

```{r}
#| label: dat1-pro-sco-t5
#| message: false

dat1_5 %>%
  ggplot(aes(proficiency, score)) +
  geom_point(alpha = 0.5) +
  geom_smooth()
```

```{r}
#| label: d1-gam-1

d1_gam_1 <- gam(
  score ~ s(proficiency),
  data = dat1_5
)

summary(d1_gam_1)
```

```{r}
#| label: d1-gam-1-check

gam.check(d1_gam_1)
```

```{r}
predict_gam(d1_gam_1, length_out = 50) %>%
  plot(series = "proficiency")
```

```{r}
d1_lm <- lm(score ~ proficiency, data = dat1)
summary(d1_lm)
```

```{r}
ggpredict(d1_lm, terms = "proficiency") %>% plot()
```

```{r}
dat1_5 %>%
  ggplot(aes(proficiency, score)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", colour = "red") +
  geom_smooth(method = "gam")
```