### 2023-12-03 scripts/02_data_analysis.qmd
edit analysis

---
title: "02 Data analysis"
format: html
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
theme_set(theme_light())
library(mgcv)
library(tidygam)
library(ggeffects)
library(dagitty)
my_seed <- 4785
```

## Read data

```{r}
#| label: data

dat1 <- readRDS("data/dat1.rds")
dat2 <- readRDS("data/dat2.rds")
```

## Data 1

### Score and proficiency (single time point)

Let's start with data from a single time point and let's model score as a function of proficiency.

```{r}
#| label: dat1-5

dat1_5 <- dat1 %>%
  filter(time == 5)

dat1_5
```

There are `r length(unique(dat1_5$subj))` subjects. The median score is `r median(dat1_5$score)` with a range between `r min(dat1_5$score)` and `r max(dat1_5$score)`. The median proficiency is `r median(dat1_5$proficiency)` with a range between `r min(dat1_5$proficiency)` and `r max(dat1_5$proficiency)`.

```{r}
#| label: dat1-pro-sco-t5
#| message: false

dat1_5 %>%
  ggplot(aes(proficiency, score)) +
  geom_point(alpha = 0.5) +
  geom_smooth()
```

```{r}
#| label: d1-gam-1

d1_gam_1 <- gam(
  score ~ s(proficiency),
  data = dat1_5
)

summary(d1_gam_1)
```

```{r}
#| label: d1-gam-1-check

gam.check(d1_gam_1)
```

```{r}
#| label: d1-gam-1-p

predict_gam(d1_gam_1, length_out = 50) %>%
  plot(series = "proficiency")
```

```{r}
#| label: d1-lm

d1_lm <- lm(score ~ proficiency, data = dat1)
summary(d1_lm)
```

```{r}
#| label: d1-lm-p

ggpredict(d1_lm, terms = "proficiency") %>% plot()
```

```{r}
#| label: d1-gam-lm

dat1_5 %>%
  ggplot(aes(proficiency, score)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", colour = "red") +
  geom_smooth(method = "gam")
```

### Score by proficiency and time

```{r}
#| label: dat1-pro-sco

dat1 %>%
  ggplot(aes(proficiency, score, group = subj)) +
  geom_path(alpha = 0.5) +
  geom_point(alpha = 0.5)
```

```{r}
#| label: d1-dag

g1 <- dagitty('dag {
proficiency [exposure,pos="-0.677,-0.105"]
score [outcome,pos="-0.070,0.385"]
time [pos="0.529,-0.098"]
proficiency -> score
time -> proficiency
}
')

exposures(g1) <- "proficiency"
outcomes(g1) <- "score"
plot(g1)
```

```{r}
#| label: d1-dag-cond

adjustmentSets(g1)
impliedConditionalIndependencies(g1)
```

```{r}
#| label: d1-gam-2

d1_gam_2 <- bam(
  score ~
    s(proficiency) + s(time) +
    s(subj, bs = "re") + s(subj, proficiency, bs = "re"),
  data = dat1
)
```

## Data 2

```{r}
#| label: dat2-plot

dat2 %>%
  ggplot(aes(timep, eud, group = id, colour = group)) +
  geom_path(alpha = 0.15, linewidth = 1) +
  facet_grid(cols = vars(vowel), rows = vars(group))
```

```{r}
#| label: prep-dat2

dat2 <- dat2 %>%
  mutate(
    group_o = ordered(group, levels = c("es", "es_0", "es_1")),
    vowel_o = ordered(vowel, levels = c("a", "i", "u")),
    group_vowel_o = ordered(interaction(group_o, vowel_o))
  )
contrasts(dat2$group_o) <- "contr.treatment"
contrasts(dat2$vowel_o) <- "contr.treatment"
contrasts(dat2$group_vowel_o) <- "contr.treatment"
```

```{r}
#| label: d2-gam

d2_gam <- bam(
  eud ~
    group_vowel_o +
    s(timep, k = 9) +
    s(timep, by = group_vowel_o, k = 9) +
    s(timep, subj, bs = "fs", m = 1) +
    s(timep, subj, by = vowel_o, bs = "fs", m = 1),
  data = dat2
)
```

```{r}
summary(d2_gam)
```

```{r}
predict_gam(
  d2_gam,
  length_out = 25,
  separate = list(group_vowel_o = c("group", "vowel")),
  exclude_terms = c("s(timep,subj)", "s(timep,subj):vowel_oi", "s(timep,subj):vowel_ou"),
  values = list(vowel_o = "a")
) %>%
  plot(series = "timep", comparison = "vowel") +
  facet_grid(cols = vars(group))
```